---
title: "Taylor Swift Song Analysis: What Makes Good and Bad Songs?"
author: "Arvind Ramakrishnan, Jason Matthew Suhari, Jason Ray, Tay Jia Jun Jeremy"
date: "`r Sys.Date()`"
output: html_document
---

## Introduction

### Background

This project analyzes Taylor Swift's song features to uncover what characteristics define a "good" or "bad" song, aiming to answer the question: "Which features make good and bad Taylor songs?" by using data from the taylor R package and Spotify's audio metrics.

### Main Question

Insert main question here

### Sub-Questions
1.) Insert sub-question #1 here

2.) Insert sub-question #2 here

3.) Insert sub-question #3 here

## Data Parsing and Initialization

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, include = TRUE,
                      fig.align = "center",  out.width = "80%")
```

### (DEBUG) Package Installations (For members only, to remove later)
```{r}
install.packages("ggplot2")
install.packages("dplyr")
install.packages("tidyverse")
install.packages("taylor")
install.packages("tidytuesdayR")
```

### Library Imports
```{r}
library(ggplot2)
library(dplyr)
library(tidyverse)
# add more here ltr
```

### Data Ingestion
```{r}
data <- tidytuesdayR::tt_load('2023-10-17')
taylor_album_songs <- data$taylor_album_songs
taylor_all_songs <- data$taylor_all_songs
taylor_albums <- data$taylor_albums
```

### Data Cleaning
```{r}

```

### Basic EDA
```{r}

```


## NEXT SECTION (CHANGE THIS HEADER)

### NEXT SECTION SUBSECTION (CHANGE THIS HEADER)

###Q2
taylor_albums %>% arrange(desc(metacritic_score),desc(user_score))

top_two_album = c("Red (Taylor's Version)", "folklore")
btm_two_album = c("reputation", "Taylor Swift")

df_plot2 <- df %>% select(album_name:tempo, user_score) %>% 
  filter(album_name %in% top_two_album | album_name %in% btm_two_album) %>%
  pivot_longer(cols= danceability:tempo, names_to = "features", values_to = "score") %>% 
  mutate(features = str_to_title(features), album_name = str_to_title(album_name))

ggplot(df_plot2, aes(x = album_name, y = score, color = album_name)) +
  geom_point(size = 2, alpha = 0.5) + 
  scale_fill_viridis(option = "viridis", discrete = TRUE) +
  facet_wrap(~ features, scales = "free") + 
  theme(axis.text.x = element_blank()) +
  labs(title = "Characteristics of Taylor's Top and Bottom 2 Albums Based on User Scores", x = "", y = "Score", color = "Album")


