---
title: "Taylor Swift Song Analysis: What Makes Good and Bad Songs?"
author: "Arvind Ramakrishnan, Jason Matthew Suhari, Jason Ray, Tay Jia Jun Jeremy"
date: "`r Sys.Date()`"
output: html_document
---

## Introduction

### Background

This project analyzes Taylor Swift's albums that were released up until 2023, with the latest being *1989 (Taylor's Version)*. 

The [dataset](https://github.com/rfordatascience/tidytuesday/blob/master/data/2023/2023-10-17/readme.md) is sourced from [TidyTuesday](https://github.com/rfordatascience/tidytuesday), featuring the following key variables:

1. **`album_name`**: Name of the album that a specific song belongs to

2. **`danceability`**: A metric of how suitable a song is for dancing (0 - 1)

3. **`energy`**: A metric of how intense and energetic a song is (0 - 1)

4. **`loudness`**: Measure of loudness in decibels (dB)

5. **`speechiness`**: A metric of how many words are spoken in a song (0 - 1)

6. **`acousticness`**: A metric of whether or not a song is acoustic, meaning void of electric amplification (0 - 1)

7. **`valence`**: A metric of how musically positive (i.e., happy, cheerful) a song is (0 - 1)

8. **`tempo`**: The overall estimated tempo of the track, measured in BPM

9. **`album_release`**: The release date of the album, meaning some songs will have the same value for this column

10. **`metacritic_score`**: A score out of 100 that represents the average critical rating for the entire album, meaning some songs will have the same value for this column

11. **`user_score`**: A score out of 100 that represents the average user rating for the entire album, meaning some songs will have the same value for this column


### Main Question

**What are the key differences, in terms of musical features, between the songs of a highly rated album and relatively lower rated album of Taylor Swift?**

### Sub-Questions
1.) Insert sub-question #1 here

2.) Insert sub-question #2 here

3.) Insert sub-question #3 here

## Data Parsing and Initialization

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, include = TRUE,
                      fig.align = "center",  out.width = "80%")
```

### (DEBUG) Package Installations (For members only, to remove later)
```{r}
# install.packages("ggplot2")
# install.packages("dplyr")
# install.packages("tidyverse")
# install.packages("taylor")
# install.packages("tidytuesdayR")
# install.packages("viridis")
```

### Library Imports
```{r}
library(ggplot2)
library(dplyr)
library(taylor)
library(tidyverse)
library(viridis)
library(lubridate)
# add more here ltr
```

### Data Ingestion & Cleaning
```{r}
data <- tidytuesdayR::tt_load('2023-10-17')
cols_to_select <- c("album_name", "danceability", "energy", "loudness", "speechiness", "acousticness", "valence", "tempo")

taylor_albums <- data$taylor_albums

df <- data$taylor_album_songs %>% 
  select(all_of(cols_to_select)) %>% 
  drop_na() %>% 
  left_join(taylor_albums, by = "album_name") %>% 
  select(-c("ep"))
```

### Exploratory Data Analysis
```{r}
glimpse(df)

# looking at distribution of all data
dist <- df %>% 
  select(where(is.numeric)) %>% 
  pivot_longer(cols = everything(), names_to = "var", values_to = "val") %>%
  ggplot(aes(x = val)) +
  geom_histogram(bins = 30) +
  labs(title = "Distribution of All Numeric Variables") +
  theme_bw() +
  facet_wrap(~var, scales = "free") +
  theme(
    plot.title = element_text(hjust = 0.5),
    axis.title.x = element_blank(),
    axis.title.y = element_blank()
    )

# finding outliers
outliers <- df %>%
  select(where(is.numeric)) %>%
  pivot_longer(cols = everything(), names_to = "var", values_to = "val") %>%
  ggplot(aes(x = var, y = val)) +
  geom_boxplot() +
  labs(title = "Boxplot for Potential Outliers in Taylor Swift's Songs") +
  theme_bw() +
  facet_wrap(~var, scales = "free") +
  theme(
    plot.title = element_text(hjust = 0.5),
    axis.title.x = element_blank(),
    axis.title.y = element_blank()
    )

heatmap <- df %>%
  select(where(is.numeric)) %>%
  cor() %>%
  as.data.frame() %>% 
  rownames_to_column(var = "x") %>%
  pivot_longer(cols = -x, names_to = "y", values_to = "value") %>%
  ggplot(aes(x = x, y = y, fill = value)) +
  geom_tile(color = "black") +
  labs(title = "Correlation Heatmap of Numerical Variables", x = "", y = "")
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1),
    plot.title = element_text(hjust = 0.5)
  ) 
```

### Visualization 1
*Line chart that plots the metacritic score as well as user score of each Taylor Swift album in the y axis against the album release date on the X axis*

**Part 1 : Data Transformation**
We will use only the `taylor_albums` dataset for this part and will remove the rows with missing values. Additionally we will scale the user scores to be weighted out of 100 by multiplying the `user_score` variable by 10 (as it is originally weighted on a scale of 1-10)
```{r}
# create a dataframe df to use for plots and analysis in this section 
df <- drop_na(taylor_albums)
# scale the user_score variable
df$user_score <- df$user_score * 10
df


```
**Part 2 : Line Chart**
Using the transformed data above that is stored in the dataframe `df`, we now plot the line chart 
```{r}
ggplot(df) + 
  geom_line(aes(x = album_release, y = metacritic_score), color = "blue", size = 2) + 
  geom_line(aes(x = album_release, y = user_score), color = "gold", size = 2) + 
  labs(title = "Metacritic and User Score of Taylor Swift Albums", 
       x = "Album Release Date", y = "Score (out of 100)") 
  
```


### Question 2
### Write question here
```{r}

```

### Question 3
Density plot that visualizes the distribution of energy in albums from the Top 2 Albums (based on metacritic and user scores) and a specific album (in this case, Taylor Swift)
```{r}
top_2_albums <- c("Red (Taylor's Version)", "folklore")

best_album_data <- df %>%
  filter(album_name %in% top_2_albums) %>%
  mutate(label = "Top 2 Albums")

album_to_compare <- "1989"
selected_album_data <- df %>%
  filter(album_name == album_to_compare) %>%
  mutate(label = album_to_compare)

data_to_plot <- bind_rows(best_album_data, selected_album_data)

ggplot(data_to_plot, 
       aes(x = energy, fill = label, color = label)) +
  geom_density(alpha = 0.5) +
  labs(title = paste("Energy Distribution of Top 2 Albums VS", album_to_compare, "Album"),
       x = "Energy",
       y = "Density") +
  theme_bw() +
  theme(
    plot.title = element_text(hjust = 0.5),
    legend.position = "bottom"
    )
```

