---
title: "Taylor Swift Song Analysis: What Makes Good and Bad Songs?"
author: "Arvind Ramakrishnan, Jason Matthew Suhari, Jason Ray, Tay Jia Jun Jeremy"
date: "`r Sys.Date()`"
output: html_document
---

## Introduction

### Background

This project analyzes Taylor Swift's song features to uncover what characteristics define a "good" or "bad" song, aiming to answer the question: "Which features make good and bad Taylor songs?" by using data from the taylor R package and Spotify's audio metrics.

### Main Question

Insert main question here

### Sub-Questions
1.) Insert sub-question #1 here

2.) Insert sub-question #2 here

3.) Insert sub-question #3 here

## Data Parsing and Initialization

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, include = TRUE,
                      fig.align = "center",  out.width = "80%")
```

### (DEBUG) Package Installations (For members only, to remove later)
```{r}
install.packages("ggplot2")
install.packages("dplyr")
install.packages("tidyverse")
install.packages("tidytuesdayR")
```

### Library Imports
```{r}
library(ggplot2)
library(dplyr)
library(tidyverse)
library(viridis)
# add more here ltr
```

### Data Ingestion
```{r}
data <- tidytuesdayR::tt_load('2023-10-17')
taylor_album_songs <- data$taylor_album_songs
taylor_all_songs <- data$taylor_all_songs
taylor_albums <- data$taylor_albums
```

### Plot 1 - which albums are good and which songs are bad?
```{r}
df1 <- drop_na(taylor_albums)

#df1
ggplot(df2) + 
  geom_line(aes(x = album_release, y = metacritic_score), color = "gold") +
  geom_line(aes(x = album_release, y = (user_score)*10), color = "blue") + 
  labs(title = "Metacritic and User Score of Taylor Swift Albums", 
       x = "Album Release Date", y = "Score (out of 100)")
  

```
### Plot 1 - Using bar plot instead of line (This is the 2nd code chunk)
```{r}
ggplot(df3) +
  geom_col(aes(x = album_name, y = score, fill = score_type), position = "dodge") +
  labs(title = "Metacritic and User Score of Taylor Swift Albums",
       x = "Album Names", y = "Score (weighted out of 100)" ) +
  scale_fill_viridis(option = "viridis", discrete = TRUE) +
  theme(axis.text.x = element_text(angle = 90)) +
  geom_text(aes(x = album_name, y = score, label = score), position = position_dodge(width = 0.9), vjust = -0.25)
```


### This is the first code chunk
```{r}
df1 <- drop_na(taylor_albums)
df2 <- df1[-c(2,4),]
df2$user_score = df2$user_score * 10
df3 <- df2 %>% pivot_longer(`metacritic_score`:`user_score`, names_to = "score_type", values_to = "score")
df3
```



## NEXT SECTION (CHANGE THIS HEADER)

### NEXT SECTION SUBSECTION (CHANGE THIS HEADER)
