---
title: "Taylor Swift Song Analysis: What Makes Good and Bad Songs?"
author: "Arvind Ramakrishnan, Jason Matthew Suhari, Jason Ray, Tay Jia Jun Jeremy"
date: "`r Sys.Date()`"
output: html_document
---

## Introduction

### Background

This project analyzes Taylor Swift's song features to uncover what characteristics define a "good" or "bad" song, aiming to answer the question: "Which features make good and bad Taylor songs?" by using data from the taylor R package and Spotify's audio metrics.

### Main Question

Insert main question here

### Sub-Questions
1.) Insert sub-question #1 here

2.) Insert sub-question #2 here

3.) Insert sub-question #3 here

## Data Parsing and Initialization

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, include = TRUE,
                      fig.align = "center",  out.width = "80%")
```

### (DEBUG) Package Installations (For members only, to remove later)
```{r}
install.packages("ggplot2")
install.packages("dplyr")
install.packages("tidyverse")
install.packages("taylor")
install.packages("tidytuesdayR")
```
#TESTING
### Library Imports
```{r}
library(ggplot2)
library(dplyr)
library(tidyverse)
# add more here ltr
```

### Data Ingestion
```{r}
data <- tidytuesdayR::tt_load('2023-10-17')
taylor_album_songs <- data$taylor_album_songs
taylor_all_songs <- data$taylor_all_songs
taylor_albums <- data$taylor_albums
```

### Data Cleaning
```{r}

```

### Basic EDA
```{r}

```


## NEXT SECTION (CHANGE THIS HEADER)

### NEXT SECTION SUBSECTION (CHANGE THIS HEADER)
