---
title: "Taylor Swift Song Analysis: What Makes Good and Bad Songs?"
author: "Arvind Ramakrishnan, Jason Matthew Suhari, Jason Ray, Tay Jia Jun Jeremy"
date: "`r Sys.Date()`"
output: html_document
---

## Introduction

### Background

This project analyzes Taylor Swift's song features to uncover what characteristics define a "good" or "bad" song, aiming to answer the question: "Which features make good and bad Taylor songs?" by using data from the taylor R package and Spotify's audio metrics.

### Main Question

Insert main question here

### Sub-Questions
1.) Insert sub-question #1 here

2.) Insert sub-question #2 here

3.) Insert sub-question #3 here

## Data Parsing and Initialization

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, include = TRUE,
                      fig.align = "center",  out.width = "80%")
```

### (DEBUG) Package Installations (For members only, to remove later)
```{r}
install.packages("ggplot2")
install.packages("dplyr")
install.packages("tidyverse")
install.packages("taylor")
install.packages("tidytuesdayR")
```

### Library Imports
```{r}
library(ggplot2)
library(dplyr)
library(tidyverse)
# add more here ltr
```

### Data Ingestion
```{r}
data <- tidytuesdayR::tt_load('2023-10-17')
taylor_album_songs <- data$taylor_album_songs
taylor_all_songs <- data$taylor_all_songs
taylor_albums <- data$taylor_albums
```

### (DEBUG) Pre-Cleaning (may or may not integrate into cleaning ltr)
```{r}
# quick data summary, a bit lengthy so I commented out for now
# str(taylor_album_songs)
# summary(taylor_album_songs)

# str(taylor_all_songs)
# summary(taylor_all_songs)

# str(taylor_albums)
# summary(taylor_albums)

# checking if we need to actually drop NA
sapply(taylor_album_songs, function(x) sum(is.na(x)))
sapply(taylor_all_songs, function(x) sum(is.na(x)))
sapply(taylor_albums, function(x) sum(is.na(x)))
# we do need to drop NA values

# checking if we need to unique / check dupe
sum(duplicated(taylor_album_songs))
sum(duplicated(taylor_all_songs))
sum(duplicated(taylor_albums))
# there are no duplicate values

# as for the rest of the data, the columns are already properly 
# named and the dates are already properly formatted. not much
# cleaning needs to be done on this part actually

### arr of var names tht we need to check for outliers
variables_to_check <- c("danceability", "energy", "loudness", "speechiness", "acousticness", "instrumentalness", "liveness", "valence", "tempo", "duration_ms")

# just change "taylor_all_songs" to whatever u wanna check!
outliers <- taylor_all_songs %>% 
                  select(variables_to_check) %>% 
                  pivot_longer(cols = everything(), names_to = "variable", values_to = "value") %>% 
  ggplot(aes(x = variable, y = value)) +
  geom_boxplot() +
  labs(title = "Box Plots of Numerical Variables in taylor_all_songs.csv ") +
  theme_minimal() +
  facet_wrap(~ variable, scales = "free") +
  theme(axis.title.x = element_blank(), axis.text.x = element_blank())

outliers

## seems there are indeed lots of outliers here and there, not sure whether or not to remove though.
```

### Data Cleaning
```{r}

```

### Basic EDA
```{r}

```


## NEXT SECTION (CHANGE THIS HEADER)

### NEXT SECTION SUBSECTION (CHANGE THIS HEADER)